---
title: "deliverable4"
output: html_document
date: "2023-03-30"
---

```{r}
library(car)
library(dplyr)
library(ggplot2)
library(hrbrthemes)
library(GGally)
library(gridExtra)
```

```{r}
daydata <- read.csv("/Users/bach_nguyen/MA 575/Labs/Project/Data/day.csv",header=TRUE)
df = data.frame(daydata)
train = df[df$yr == '0',]
attach(train)
```

Matrix plot:

```{r}
data <- data.frame(casual, temp, atemp, hum, windspeed)
ggpairs(data, upper = list(continuous = wrap("points", alpha = 0.3,    size=0.1)),
lower = list(continuous = wrap('cor', size = 4)))
```

```{r}
season_f = as.factor(season)
workingday_f = as.factor(workingday)
holiday_f = as.factor(holiday)
weekday_f = as.factor(weekday)
weathersit_f = as.factor(weathersit)
par(mfrow=c(2,2))
boxplot(casual~season_f,ylab="casual", xlab="season")
boxplot(casual~workingday_f,ylab="casual", xlab="workingday")
boxplot(casual~weathersit_f,ylab="casual", xlab="weathersit")
boxplot(casual~weekday_f,ylab="casual", xlab="weekday")
```

\--\> Holiday or non-working day get higher means.

```{r}
library(MASS)
pairs(log(casual)~log(temp)+log(atemp) + log(hum) + log(windspeed),data=daydata,gap=0.4,cex.labels=1.5)
```

First try: try to plug all variables in

```{r}
m1_full.mls = lm(casual ~ temp +atemp + hum + windspeed + season_f + workingday_f + weathersit_f + weekday_f)
summary(m1_full.mls)
```

```{r}
vif(m1_full.mls)
```

From the multi collinearity plot, we can see temp and atemp are highly correlated.

\-\-\-\-\-\-\-\-\-\-\-\--

Second try: delete atemp, and holiday_f.

```{r}
ggplot(data = train) + geom_point(data=train, aes(x=1:nrow(data), y=casual, color=workingday_f)) 
```

-\>This plot shows that non-working day yields more bike rents.

```{r}
m2_full.mls = lm(casual ~ temp + hum + windspeed + season_f +weekday_f + workingday_f:(temp + hum + windspeed + + season_f) )
summary(m2_full.mls)
```

```{r}
m2_reduced.mls = lm(casual ~ temp + hum + windspeed )
summary(m2_reduced.mls)
```

```{r}
anova(m2_reduced.mls, m2_full.mls)
```

-\> Should use the full model with all the categorical variables.

\-\-\-\-\-\-\-\-\-\-\--

Begin the diagnostic:

```{r}
StanRes <- rstandard(m2_full.mls)
par(mfrow=c(3,3))
plot(temp,StanRes, ylab="Standardized Residuals")
plot(hum,StanRes, ylab="Standardized Residuals")
plot(windspeed,StanRes, ylab="Standardized Residuals")
boxplot(StanRes~season_f,ylab="Standardized Residuals",xlab="season_f")
boxplot(StanRes~workingday_f,ylab="Standardized Residuals",xlab="workingday_f")
boxplot(StanRes~weathersit_f,ylab="Standardized Residuals",xlab="weathersit_f")
boxplot(StanRes~weekday_f,ylab="Standardized Residuals",xlab="weekday_f")
```

```{r}
par(mfrow=c(1,1))
plot(m2_full.mls$fitted.values,casual,xlab="Fitted Values")
abline(lsfit(m2_full.mls$fitted.values,casual))
plot(m2_full.mls$fitted.values,log(casual),xlab="Fitted Values")
abline(lsfit(m2_full.mls$fitted.values,log(casual)))
```

```{r}
par(mfrow=c(2,2))
plot(m2_full.mls)
abline(v=2*8/72,lty=2)
```

-\> need to check point 442,551,725

```{r}
vif(m2_full.mls)
```

```{r}

par(mfrow=c(2,2))
mmp(m2_full.mls,temp)
mmp(m2_full.mls,hum)
mmp(m2_full.mls,windspeed)
mmp(m2_full.mls,m2_full.mls$fitted.values,xlab="Fitted Values")
```

```{r}
library(car)
par(mfrow=c(2,4))
avPlot(m2_full.mls,variable=temp,ask=FALSE, main="")
avPlot(m2_full.mls,variable=hum,ask=FALSE, main="")
avPlot(m2_full.mls,variable=windspeed,ask=FALSE, main="")

```

```{r}
avPlots(m2_full.mls,variable=season_f,ask=FALSE, main="")
avPlots(m2_full.mls,variable=workingday_f,ask=FALSE, main="")
avPlots(m2_full.mls,variable=weathersit_f,ask=FALSE, main="")
```

\-\-\-\-\-\-\-\-\-\-\-\-\--

Validation:

```{r}
validation = df[df$yr == 1,]
```

```{r}
# Residuals for training data
ResMLS <- resid(m2_full.mls)

# Mean Square Error for training data
mean((ResMLS)^2)
```
