---
title: "4_3"
output:
  pdf_document: default
  html_document: default
date: "2023-03-30"
---

```{r}
library(car)
library(dplyr)
library(ggplot2)
library(hrbrthemes)
library(GGally)
library(gridExtra)
```

1.  Model selection

```{r}
daydata <- read.csv("/Users/bach_nguyen/MA 575/Labs/Project/Data/day.csv",header=TRUE)
df = data.frame(daydata)
train = df[df$yr == '0',]
validation = df[df$yr == '1',]
attach(train)
```

```{r}
data <- data.frame(casual, temp, atemp, hum, windspeed)
ggpairs(data, upper = list(continuous = wrap("points", alpha = 0.3,    size=0.1)),
lower = list(continuous = wrap('cor', size = 4)))
```

```{r}
par(mfrow=c(2,2))
boxplot(casual~as.factor(season),ylab="casual", xlab="season")
boxplot(casual~as.factor(workingday),ylab="casual", xlab="workingday")
boxplot(casual~as.factor(weathersit),ylab="casual", xlab="weathersit")
boxplot(casual~as.factor(weekday),ylab="casual", xlab="weekday")
```

first try:

```{r}
m1 = lm(casual ~ temp + windspeed)
summary(m1)
```

```{r}
vif(m1)
```

-\> From the multi collinearity plot, we can see temp and atemp are highly correlated. And temp is more significant, keep temp

```{r}
summary(aov(casual~season, train))
```

-\> season is significant to add to the model

```{r}
summary(aov(casual~workingday, train))
```

-\> workingday is significant to add to the model

```{r}
summary(aov(casual~weekday, train))
```

-\> weekday is not significant

```{r}
summary(aov(casual~holiday, train))
```

-\>holiday is not significant

from the boxplot and below scatter plot, non-working days have higher bike rents.

```{r}
ggplot(data = train) + geom_point(data=train, aes(x=1:nrow(data), y=casual, color=as.factor(workingday))) 
```

```{r}
plot(m1$fitted.values,casual,xlab="Fitted Values")
abline(lsfit(m1$fitted.values,casual))
```

-\> Use log(casual)

```{r}
plot(m1$fitted.values,log(casual),xlab="Fitted Values")
abline(lsfit(m1$fitted.values,log(casual)))
```

-\> fit better

2.  Choosing MLR

```{r}
m = lm(log(casual) ~ temp + I(temp^2) + windspeed + as.factor(season)+ as.factor(weathersit)+ as.factor(workingday))
summary(m)
```

```{r}
m1_with_log <- lm(log(casual) ~ temp + hum + windspeed)
anova(m1_with_log, m)
```

-\> there is statistically evidence to use full model.

Begin the diagnostics:

```{r}
StanRes <- rstandard(m)
par(mfrow=c(3,3))
plot(temp,StanRes, ylab="Standardized Residuals")
plot(hum,StanRes, ylab="Standardized Residuals")
plot(windspeed,StanRes, ylab="Standardized Residuals")
boxplot(StanRes~as.factor(season),ylab="Standardized Residuals",xlab="season")
boxplot(StanRes~as.factor(workingday),ylab="Standardized Residuals",xlab="workingday")
boxplot(StanRes~as.factor(weathersit),ylab="Standardized Residuals",xlab="weathersit")
```

```{r}
plot(m$fitted.values,log(casual),xlab="Fitted Values")
abline(lsfit(m$fitted.values,log(casual)))
```

```{r}
par(mfrow=c(2,2))
plot(m)
abline(v=2*8/72,lty=2)
```

```{r}
vif(m)
```

```{r}
par(mfrow=c(2,2))
mmp(m,temp)
mmp(m,hum)
mmp(m,windspeed)
mmp(m,m$fitted.values,xlab="Fitted Values")
```

```{r}
library(car)
par(mfrow=c(2,2))
avPlot(m,variable=temp,ask=FALSE, main="")
avPlot(m,variable=hum,ask=FALSE, main="")
avPlot(m,variable=windspeed,ask=FALSE, main="")
```

\-\-\-\-\-\-\-\-\-\-\-\-\-\--

Validation

```{r}
# Residuals for training data
ResMLS <- resid(m)

# Mean Square Error for training data
mean((ResMLS)^2)
```

```{r}
# Mean Square Error for validation data

# Residuals for validation data
#If the logical se. fit is TRUE , standard errors of the predictions are also calculated.
new_data <- data.frame(temp=validation$temp,hum=validation$hum, windspeed =validation$windspeed, season =validation$season, weathersit=validation$weathersit, workingday= validation$workingday)

output <- predict(m,se.fit = TRUE, newdata = new_data)

ResMLSValidation <- log(validation$casual) - output$fit

mean((ResMLSValidation)^2)

#mean((output$residual.scale)^2)
```

```{r}
# Relative Mean Square Error for validation data
mean((ResMLSValidation)^2) / mean((validation$casual)^2)

#mean((output$residual.scale)^2) / mean((validation$casual)^2)
```

```{r}
plot(output$fit,validation$casual,xlab="prediction")
abline(lsfit(output$fit,validation$casual))
```

```{r}
# Create data frame with validation observation and prediction
test = data.frame(validation$casual,exp(output$fit), 1:length(output$fit));
colnames(test)[1] = "Casual"
colnames(test)[2] = "Prediction"
colnames(test)[3] = "Index"

```

```{r}
ggplot(data = test, aes(x = Index)) +
  geom_line(aes(y = Casual, color = "Casual")) + 
  geom_line(aes(y = Prediction, color="Prediction"), linetype="twodash") +  
  scale_color_manual(name = element_blank(), labels = c("Casual","Prediction"),
                     values = c("darkred", "steelblue")) + labs(y = "") + 
  ggtitle("Validation")
```
